\chaptermark{Acknowledgements}
\begin{center}
\textbf{Acknowledgements}
\end{center}

My time at Brown has been some of the most rewarding years of my life, both professionally and personally.
This thesis presents the culmination of my work here, but the lessons I have learnt go beyond this.
I am deeply thankful for being given this opportunity, and I have many people to thank for making this time so special.\par

First, I would like to thank my advisor Professor Eric Suuberg.
He was the first to interest me in vapor intrusion research, and without his support I would not have been able to pursue this research at such a fine institution.
Throughout my years at Brown, Eric has always been available to guide and mentor me, and his encouragement has been instrumental for my development.
It has been a privilege to have worked with him.\par

Secondly, I would like to thank the other two committee members - Professors Brenda Rubenstein and Franklin Goldsmith.
I have greatly benefited from their commitment to teaching excellence and I have thoroughly enjoyed attending their classes.
I am deeply grateful for their service on this committee.\par

I also want to offer my thanks to the many professors I have met at Brown and whose classes I have attended - it truly has been a wonderful experience.
I would especially like to thank Indrek Külaots and Andrew Peterson for giving me the opportunity to assist them in teaching their classes.
Through them I have discovered a passion for teaching that I did not think that I had.\par

To all of my friends and colleagues at Brown:
Thank you for your support and for being in my life.
Your companionship has contributed to my enjoyment and well-being in ways I cannot express.
I would especially like to thank Zachary Saleeba, Jordan Noble, Aaron North, and Adam Gavarkovs for being such wonderful friends.\par

Lastly, I would like to thank my parents and my family.
Moving and living in another country has not easy, and your tireless support has been instrumental for my success.
I am especially thankful that my father, Martin Ström, has been able to visit me so regularly - I have many fond memories of our times together.
I love you all with all of my heart.\par

\vfill
\newpage
