\chapter{Geometry Generation}

To create our quarter geometry, only a few simple geometric objects and Boolean operations are required: two cuboids, two rectangles, one Boolean difference operation, and one Boolean join operation.
Figure \ref{fig:geometry} shows the resulting geometry.
Note that $z = \SI{0}{\metre}$ is the groundwater/soil interface and the plane of symmetry is around the $(x, y) = (\SI{0}{\metre},\SI{0}{\metre})$ axis\par

To create the soil surrounding the building using the COMSOL geometry generator:
\begin{enumerate}
  \item Create a \SI{15}{\metre} by \SI{15}{\metre} by \SI{4}{\metre} block with its base at $(x, y, z) = (\SI{0}{\metre},\SI{0}{\metre},\SI{0}{\metre})$. This is the entire soil domain.
  \item Create a \SI{5}{\metre} by \SI{5}{\metre} by \SI{1}{\metre} block with its base at $(x, y, z) = (\SI{0}{\metre},\SI{0}{\metre},\SI{3}{\metre})$. This will represent the volume that the house take up in the soil, i.e. the underground portion of the basement.
  \item Perform a difference operation, removing the "basement" block from the "soil" block.
\end{enumerate}
At this point you will see that a quarter soil domain has been created, with an empty space that represents a house with a foundation slab located \SI{1}{\metre} bgs.\par

The foundation crack will be modeled by joining two  \SI{1}{\centi\metre} wide strip that spans the perimeter of the surface that represents the house foundation.
This strip is created by joining two rectangles on foundation surface:
\begin{enumerate}
  \item Define a work plane \SI{3}{\metre} above zero. This allows us to place two-dimensional objects on the surface of or inside a three-dimensional object.
  \item On the work plane create a \SI{5}{\metre} by \SI{1}{\centi\metre} rectangle with its base at $(x, y) = (\SI{0}{\metre},\SI{5}{\metre} - \SI{1}{\centi\metre})$. This represents one side of the perimeter crack.
  \item Copy the rectangle and rotate it \SI{90}{\degree} around the corner of the foundation, i.e. $(x, y) = (\SI{5}{\metre} - \SI{0.5}{\centi\metre},\SI{5}{\metre} - \SI{0.5}{\centi\metre})$.
  \item Join the two rectangles to create a unified perimeter foundation crack.
\end{enumerate}
Now the geometry of this VI scenario is complete.\par
